---
title: "Florida Palmetto (*Sabal etonia* and *Serenoa repens*): 1981-2017"
author: "Genevieve Chiong"
date: "1/31/2021"
output:
   html_document:
     theme: cosmo
     code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

library(tidyverse)
library(GGally)
library(broom)
library(here)
library(kableExtra)
library(janitor)
```


```{r}
# Read in the data
palmetto <- read_csv(here("data", "palmetto.csv")) 

# Add species names and select for variables of interest
palmetto_clean <- palmetto %>%
  mutate(species_name = case_when(species == 1 ~ "Serenoa repens",
                                  species == 2 ~ "Sabal etonia")) %>% 
  select(species_name, height:green_lvs)
  
# Change species_name column to a factor 
palmetto_clean$species_name <- as.factor(palmetto_clean$species_name)
```

#### **Data visualizations exploring differences in plant height, canopy length, canopy width, and number of green leaves for two Florida palmetto species, *Sabal etonia* and *Serenoa repens***
```{r}
# Take a look at the data
#palmetto_clean %>% 
  #ggpairs(aes(color = species_name)) 

# Check the levels (Sabal etonia is 0, Serenoa repens is 1)
#levels(palmetto_clean$species_name) 
```

```{r, fig.cap = "**Figure 1.** Comparison of plant height (cm) and canopy length (cm) between two Florida palmetto species, *Sabal etonia* (teal) and *Serenoa repens* (magenta), using data from 1981-2017. Data: Abrahamson, W.G. 2019."}

# Look at trends across variables for the two palmetto species 

# Graph of height vs. canopy length

ggplot(data = palmetto_clean, 
       aes(x = height, 
           y = length)) +
  geom_jitter(aes(color = species_name)) + 
                scale_color_manual(values=c("cadetblue4", "mediumvioletred")) +
                labs(x = "\nHeight (cm)",
                     y = "Canopy Length (cm)\n", 
                     color = "Species") +
                theme_minimal() 
```

*Sabal etonia* and *Serenoa erpens* have overall similar plant height (cm) and canopy length (cm). However, *Sabal etonia* has a slightly larger canopy length and *Serenoa repens* has a slightly greater height. 

```{r, fig.cap = "**Figure 2.** Comparison of canopy width (cm) and number of green leaves between two Florida palmetto species, *Sabal etonia* (teal) and *Serenoa repens* (magenta), using data from 1981-2017. Data: Abrahamson, W.G. 2019."}

# Graph of canopy width vs number of green leaves
   
   ggplot(data = palmetto_clean, 
       aes(x = width, 
           y = green_lvs)) +
 geom_jitter(aes(color = species_name)) + 
  scale_color_manual(values=c("cadetblue4", "mediumvioletred")) +
  labs(x = "\nCanopy Width (cm)",
       y = "Green Leaves Count\n", 
       color = "Species") +
  theme_minimal() 
```

*Sabal etonia* has a slightly larger canopy width (cm) than *Serenoa repens*, whereas *Serenoa repens* has considerably more green leaves. 


#### **Binary Logistic Regression**

```{r}
# Binary logistic regression using plant height, canopy length, canopy width, and number of green leaves as predictor values to understand how they relate to the probability of a plant being Serenoa repens or Sabal etonia

# Sabal etonia is first '0' factor level, Serenoa repens is '1'

palmetto_blr <- glm(species_name ~ height + length + width + green_lvs, 
                    data = palmetto_clean,
                    family = "binomial")

# Look at the model
#palmetto_blr

#summary(palmetto_blr)

# Get a tidy version w/broom
palmetto_blr_tidy <- broom::tidy(palmetto_blr)
  
# Coefficients are for the predictor variables that relate to the log odds of the "Serenoa repens" outcome (level = 1)

#summary(palmetto_blr_tidy)
```

**Table 1.** The binary logistic regression model results using plant height, canopy length, canopy width, and number of green leaves as predictor variables that relate to the probability of a plant being *Serenoa repens*. Data: Abrahamson, W.G. 2019.

```{r}
# Create a table containing the binary logistic regression model results

palmetto_blr_tidy$p.value[palmetto_blr_tidy$p.value < 0.0001] <- "<0.0001"

kable(palmetto_blr_tidy,
      col.names = c("Term",
                    "Estimate",
                    "Standard Error",
                    "Statistic",
                    "P-value")) %>% 
  kable_styling(bootstrap_options = "striped",
                full_width = FALSE) 
```


#### **Classification success rate of binary logistic regression model** 

**Table 2.** The number of *Serenoa repens* and *Sabal etonia* palmetto plants correctly and incorrectly classified using the binary logistic regression model with four predictor variables: plant height, canopy length, canopy width, and number of green leaves. Data: Abrahamson, W.G. 2019.

```{r}
# Evaluate how successfully the binary logistic regression model would "classify" a plant as the correct species, using a 50% cutoff 

blr_fitted <- palmetto_blr %>% 
  broom::augment(type.predict = "response")

blr_fitted_success <- blr_fitted %>% 
  mutate(predicted_species = case_when(.fitted >= 0.5 ~ "Serenoa repens",
                                       .fitted < 0.5 ~ "Sabal etonia")) 

blr_fitted_success$correct_classification <- ifelse(blr_fitted_success$species_name == blr_fitted_success$predicted_species, "Correct", "Incorrect")

# Create a table that shows the count of plants correctly classified and incorrectly classified
palmetto_counts <- blr_fitted_success %>% 
  tabyl(species_name, correct_classification) 

# Add column for % correctly classified 
palmetto_counts_edited <- palmetto_counts %>% 
  mutate(percent_correct = round((palmetto_counts$Correct/(palmetto_counts$Correct + palmetto_counts$Incorrect)*100), 2))

# Create a finalized table for count of plants correctly and incorrectly classified as well as % correctly classified
kable(palmetto_counts_edited, 
      col.names = c("Species",
                    "Correctly Classified",
                    "Incorrectly Classified",
                    "% Correctly Classified")) %>% 
  kable_styling(bootstrap_options = "striped",
                full_width = FALSE)
```


#### **Citations**

Abrahamson, W.G. 2019. Survival, growth and biomass estimates of two dominant palmetto species of south-central Florida from 1981 - 2017, ongoing at 5-year intervals ver 1. Environmental Data Initiative. https://doi.org/10.6073/pasta/f2f96ec76fbbd4b9db431c79a770c4d5






