---
title: "Florida Palmetto"
author: "Genevieve Chiong"
date: "1/31/2021"
output:
   html_document:
     code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

library(tidyverse)
library(GGally)
library(broom)
library(jtools)
library(here)
```


Data source: Abrahamson, W.G. 2019. Survival, growth and biomass estimates of two dominant palmetto species of south-central Florida from 1981 - 2017, ongoing at 5-year intervals ver 1. Environmental Data Initiative. https://doi.org/10.6073/pasta/f2f96ec76fbbd4b9db431c79a770c4d5

```{r}
# Read in the data
palmetto <- read_csv(here("data", "palmetto.csv")) %>% 
  mutate(species_name = case_when(species == 1 ~ "Serenoa repens",
                                  species == 2 ~ "Sabal etonia")) %>% 
  select(species_name, height:green_lvs)
  

palmetto$species_name <- as.factor(palmetto$species_name)
```

### Data visualizations exploring differences in height, canopy length, canopy width, and green leaves for Sabal etonia and Serenoa repens
```{r}
palmetto %>% # Take a look at the data 
  ggpairs(aes(color = species_name)) 

levels(palmetto$species_name) # Check the levels (Sabal etonia is 0, Serenoa repens is 1)
```

```{r, fig.cap = "**Figure 1.** Comparison of plant height (cm) and canopy length (cm) between two Florida palmetto species, Sabal etonia (teal) and Serenoa repens (magenta), using data from 1981-2017. Data: Abrahamson, W.G. 2019"}

# Look at trends across variables for the two palmetto species 

# Graph of height vs. canopy length
ggplot(data = palmetto, 
       aes(x = height, 
           y = length)) +
  geom_point(aes(color = species_name)) +
  facet_wrap(~species_name) +
  scale_color_manual(values=c("cadetblue4", "mediumvioletred")) +
  labs(x = "\nHeight (cm)",
       y = "Canopy Length (cm)\n", 
       color = "Species") +
  theme_minimal() 
```

Sabal etonia and Serenoa erpens have overall similar plant height (cm) and canopy length (cm). However, Sabal etonia has a slightly larger canopy length and Serenoa repens has a slightly greater height. 

```{r, fig.cap = "**Figure 2.** Comparison of canopy width (cm) and green leaves count between two Florida palmetto species, Sabal etonia (teal) and Serenoa repens (magenta), using data from 1981-2017. Data: Abrahamson, W.G. 2019."}

# Graph of canopy width vs green leaves count 
ggplot(data = palmetto, 
       aes(x = width, 
           y = green_lvs)) +
  geom_point(aes(color = species_name)) +
  facet_wrap(~species_name) +
  scale_color_manual(values=c("cadetblue4", "mediumvioletred")) +
  labs(x = "\nCanopy Width (cm)",
       y = "Green Leaves Count\n", 
       color = "Species") +
  theme_minimal() 
```

Sabal etonia has a slightly larger canopy width (cm) than Serenoa repens, whereas Serenoa repens has considerably more green leaves. 

### Binary Logistic Regression 





The output should be a finalized table containing the binary logistic regression model results (at least coefficients, standard errors for the coefficients, and information for significance - consider using broom::tidy() to get you most of the way). 








